Busch, Khakhalin. Intrinsic temporal tuning

### 2019-05-31 reviews and responses.md

> Busch and Khakhalin use dynamic clamp to generate synaptic conductances of different durations and
then measure the number of action potentials the neurons fire in response to the different
conductances. Using this approach it is found that intrinsic excitability of tectal neurons is temporally
tuned – meaning that they fire best in response to certain durations of synaptic input. This finding is
interesting and unique because typically intrinsic excitability is quantified by generating “input-output”
curves which only vary the amplitude of the synaptic drive, not the duration. Furthermore, the authors
find that various visual stimuli lead to an “adaptive” – not homeostatic – type of temporal tuning of
intrinsic properties such that neurons exposed to shorter stimuli became more responsive to shorter
stimuli and less responsive to longer stimuli., and visa versa. In this study, the effects of visual and
auditory conditioning (and combos thereof) on temporal tuning of both synaptic and intrinsic
excitability are determined. Overall, this study makes great use of dynamic clamp. The interpretation of
the results is limited because the ways in which the conditioning stimuli actually activate tectal neurons
is not empirically established. Clarification, simplification, and consistent use of terms would improve
this paper.

### Minor

> Page 5, line 45/46: “As one possible explanation, auditory and mechanosensory inputs may have
different cellular or subcellular 46 targets in the tectum (Bollmann and Engert, 2009).”
This Bollmann and Engert reference does not address mechanosensory and auditory inputs. This paper
shows that the RGC inputs form a topographic map along the proximal-distal axis of individual tectal
neuron dendrites. This reference should be replaced with an appropriate reference.

We replaced this sentence with the following sentence: "This was not necessarily surprising, as auditory and mechanosensory inputs are known to have different cellular targets in the tectum (Pratt and Aizenman, 2009; Felch et al., 2016; Truszkowski et al., 2017), and thus can recruit tectal excitatory and inhibitory networks to a different extent."

@@

> Also, how can having different subcellular targets could explain the results? This concept needs to be
clarified – an example would be helpful.

We now added the following sentence to the text: "Alternatively, acoustic stimuli may be inherently weaker in triggering plasticity effects in tectal neurons, as compared to visual inputs, they arrive at different subcellular compartments within the dendritic tree (Hiramoto and Cline, 2009; Deeg et al., 2009), which can affect their influence on neuronal plasticity (Richards and Lillicrap, 2019)."

The actual rationale here is that inputs to different parts of the dentritic tree can trigger action potentials of different shape, and differentially affect calcium dynamics in the neuron, which, together with non-linear processing in dendrites may make different types of inputs unequal in terms of their effects on neuronal activity at one hand, and their impact on their tuning on the other. Two extreme examples of this effect would be the impact of climbing fibers on the activity of Purkinje cells in the cerebellum, and inputs to the apical dendrites of pyramidal cells in the cortex. However, we feel that this topic is really deep, yet somewhat tangential to the main message of the paper, so we tried to limit our exposition to this short sentence with a reference.

@@

> Page 13 line 46: The authors use the term “distributed sensory network” . Is a “distributed” network a
special type of sensory network?

Technically, even two neurons form a network (a true mathematician would probably argue that even one neuron forms a network!), so we tried to remind the reader that the tectal network is a relatively large one, with broad recurrent excitation, and in which different parts of the sensory input are processed in parallel. Thus we used the word "distributed", as in "distributed computations", as opposed to more local and covergent networks, such as, for example, the Mauthner cell network in the hindbrain. If you think that this word is not appropriate here, please let us know, and we'll remove it.

@@

> Page 3, line 26,27,28: “In the tadpole tectum, looming stimuli are known to elicit stronger responses
compared to instantaneous inversions, yet the suppression yielded by looming stimuli was weaker”.
The logic of this sentence is not clear. Why exactly would the authors not expect that the looming
stimuli that is known to elicit stronger (synaptic?) responses would also cause less suppression (of action
potentials?) than does flashes?

This was unexpected as intrinsic plasticity is typically described as a homeostatic mechanism, and thus a stronger stimulus should, in principle, cause stronger suppression of spiking. To make it more clear, we now changed this sentence:

"This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses (Khakhalin et al., 2014; Khakhalin, 2019), yet their effect on neuronal excitability was weaker, seemingly violating the principle of homeostatic plasticity (Pratt and Aizenman, 2007; Turrigiano, 2007)."

@@

### Major

> Page 5, line 2,3,4 : But do the authors actually know how/if their conditioning stimuli is altering the
pattern/strength of synaptic input? If so, this data should be shown. If not, it should be made clear that
this is an assumption/ theoretical prediction.

In the submitted document, these lines said:

"We then examined whether different types of sensory activation would differentially reshape  
temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed  
homeostatic (increased activation led to reduced spiking), by the same logic, one could
expect shorter stimuli (flashes) to selectively suppress responses to shorter synaptic inputs."

Here we are not talking about our conditional stimuli alterning the pattern of synaptic inputs; we just refer to the fact that during the stimulation, shorter stimuli (such as instantaneous flashes, that are technically the shortest stimuli possible) would cause faster visual inputs to the tectum, compared to longer stimuli (second-long continuous transitions) that would cause slower inputs. In this study, we did not measure actual synaptic responses in the tectum during conditioning, but from previous studies we know that responses to flashes are actually much shorter than responses to looming stimuli (Khakhalin 2014), as described in the beginning of the "Results" section, and then again in the "Methods".

We now changed the wording of this sentence, to make this logic clearer.

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; Khakhalin et al. 2014) would selectively suppress spiking in repsonse to shorter synaptic inputs."

@@

> Page 5, line 5: “As changes in intrinsic properties seemed homeostatic (increased activation led to
reduced spiking), by the same logic, one could expect shorter stimuli (flashes) to selectively suppress
responses (action potential firing or synaptic responses or both?) to shorter synaptic inputs”.
What intrinsic properties were measured? It appears that only number of spikes was counted. The
authors report that long-term increase in activation (by visual stimuli) decreased spiking – but was that
due to a decrease in intrinsic excitability or because the stimuli decreased synaptic strengths – which in
turn led to the observed decrease in number of spikes fired? This is critical to define.
Second, do the authors mean to say that shorter stimuli would actually (somehow) suppress responses?
How would that work?

All our measurements of the number of spikes relied on the dynamic clamp experiments, that used simulated synaptic inputs, and not real synaptic stimulation. It means that any changes in spiking we observed in the previous section of the results is necessarily coming from the change in intrinsic properties. While a change in the synaptic strength in response to sensory stimulation is possible, and indeed is investigated below in this paper, it would not have affected the number of spikes we reported in figures 1 and 2, as they were recorded in dynamic clamp experiments.

As described above, to make this point more clear, we now reworded this sentence:

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; Khakhalin et al. 2014) would selectively suppress spiking in repsonse to shorter synaptic inputs."

@@

> Page 5 line 15: Does the word “preferred” (in quotes) mean that the cells fired more action potentials?
Maybe refrain from personification.

We hope that the word "prefer" is intuitive enough to most readers, and it is also way shorter than "responded with more spikes", or "produced more action potentials", which allows for easier writing, and hopefully better reading. Realistically, we believe that most readers of this article will not be at risk of truly personifying neurons, so while we take care to never personify animals, we are intentionally a bit more relaxed when writing about cells, channels, receptors etc. Still, to make this sentence better, we removed the quotes around "preferred", to make it more obvious that it is not an antropomorphic allegory, but just a normal word use. Overall, we use the word "prefer" 6 times in the paper, always in the same way, when comparing spiking responses to different stimuli.

@@

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. In most cases, it may be expected to underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). For each particular comparison, however, both d, F and p values would differ by a small random value. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (`smean~rostral+medial+group`) produces a p-value of 0.01. Without adjustment (`smean~group`) we get a p-value of 0.03. For temporal tuning, correct analysis (`sbend~rostral+medial+group`) gives p=0.0304, and if covariate is left unadjusted (`sbend~group`) we get p=0.029. Similarly, for amplitude tuning, we get p=0.00209 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, the optimal models were not qualitatively different between adjusted and unadjusted data, but all values in the table were slightly different.

If you would like us to numerically re-run any part of the analysis without a position adjustment, please let us know which part you would like to see. All scripts that produce technical checks, visualizations, and analyses related to this adjustment are available for inspection at: 
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/dynamic_other_vars.R

@@

> Page 13, line 7: What does “intrinsic spikiness” mean? Is this another term for intrinsic excitability, or is
this a different parameter? In fact, throughout the manuscript, there are many terms including “intrinsic
excitability, spikiness, number of spikes, and intrinsic spikiness that probably mean the same thing – or
maybe not. If these terms all mean the same thing, then one term should be chosen and used
throughout. If they all refer to different entities, then each one should be clearly defined.

In this updated version, we consistently use "spikiness" as a name for the variable that is for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

The expression "Intrinsic spikiness" is no longer used in the text.

The expression "Intrinsic excitability" is now used only as a name for the general phenomenon, that technically may be probed with different alternative protocols and measures. In this  study, we mostly assess it using a more narrowly defined "spikness" variable, and then, towards the end, using a less informative measure of the number of spikes observed in current clamp experiments. In the text, we now consistely use "intrinsic excitability" when we make inferences about the actual properties of the neuron, and use "spikiness" when we talk about the actual value we calculated to assess and quantify intrinsic excitability.

@@

> - In reference to the text at the bottom of page10: Is it possible to include a summary
figure of data from the stepwise generalized linear model in Fig. 4? The purpose of
this part of the study was to try and identify electrophysiological mechanism(s) that
explain the tuning variables. The heart of Fig. 4 should be that they don’t, and figure
that captures this would be helpful.

We are not sure how this fact can be illustrated. We searched for papers that use stepwise linear models, and it seems they do not usually include summary figures to describe these models. (They often include tables, listing R-squared, F, df, and p-values for every component, but as in this paper we are not particularly concerned by the precise composition of our models, we doubt a table like that would be useful here). Do you have any specific type of illustration in mind, in which case, could you provide a reference, or maybe describe it? The only type of illustration we could come up with is a scatterplot that compares values predicted from the model to actual values, but it is already shown in the paper, as Figure 4E.

> - If these data are interpreted as homeostatic (neurons became less excitable
following sensory experience), is there any evidence that tectal neurons have
different firing activities during the different kinds of sensory experience? For
example, does sound engage tectal neurons less than visual stimuli, and might this
explain some of the trends in the data?

This may indeed explain some of the trends in the data, but unfortately, as we described above, recording of tadpole tectal reponses to sounds does not seem to be possible in our lab, and is fundamentally technically problematic in principle. We can only guess that stimuli so different as flashes, looms, and sound clicks, would produce different responses in the tectum, as their dynamics is drastically different, as they engage different circuits before they rich the tectum (retinal for visual stimuli, hindbrain for acoustic), which undoubtedly shapes the dynamics of tectal activation (based on Khakhalin 2014, Felch 2016, Truszkowski 2017, Khakhalin 2019).

We now make this point more clear by referencing it in the beginning of the paper (as described above), and then again, at the very end of Discussion, by describing how our observations can be used to predict actual, and yet unobserved, activation of the tectum during multisensory stimulation in vivo:

"Finally, based on the multisensory phenomena reported in this paper, we predict that even though multisensory stimulation tends to increase tectal responses in vitro (Felch et al., 2016; Truszkowski et al., 2017), it would be likely to reduce peak activation in vivo."
